A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution

### 2013-03-12 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex

First of all, the authors wish to express their gratitude to the reviewers for their encouraging comments and constructive criticisms, which we tried to address in this revised version of our manuscript.

\section{Detailed responses}

\subsection{Reviewer \#1}

\Comment{The benchmark only compares to RNAinverse which is outperformed by most methods you name in your introduction. Thus, a real benchmarking would have to compare to other methods as well. For most, source code packages are available.}
\Answer{In addition to \RNAinverse~\cite{Hofacker:1994}, when possible, we systematically benchmarked \ourprog against an extended selection of available tools, including \NUPACK~\cite{Zadeh:2011fk}, \frankenstein~\cite{Lyngso:2012vn}, \RNASSD~\cite{Aguirre-Hernandez:2007kx}, \INFORNA~\cite{Busch:2006uq}. Technical difficulties prevented us to include \RNAexinv~\cite{Avihoo:2011fk} in this draft, but the latter could be easily included later. While the page limitations associated with a conference submission do not allow for a an exhaustive discussion, we now report in the revised manuscript some of the most striking results. Mainly, with the exception of \RNASSD, none of the latest version of our competitor allows an explicit control over the \GCContent, resulting in designs that are usually highly concentrated in \GCContent (with means ranging from 50\% to 80\% depending on the software).
At low {\GCContent}s, our results showed that \ourprog is more efficient (i.e. fast and scalable to large structure) than local search approaches and generate more diverse sequences.
}

### 2013-03-12 ISMB/results_ISMB.tex
more typos

Figure~\ref{fig:time} presents the average times spent running \ourprog and \RNAinverse to generate one sequence
with the required \GCContent. As expected, the time grows linearly
in function of the length of the structures for \ourprog.  A highly intriguing feature is the substantial decrease of the time 
required for running \RNAinverse when $15\%$ of nucleotides are unpaired in the target structure.
However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).

@@

As discussed in Sec.~\ref{sec:implementation}, we noticed a highly decreased computational time needed to generate the sequences with $15\%$ free 
nucleotides in the loops. We remark that those structures also yield a much lower structural sensitivity.

@@

The sequence identity is defined over a set $\mathcal{S}$ of aligned sequences (in our case, all sequences have the same length and can be trivially aligned) as :
\begin{align}
  \sum_{s^1,s^2 \in\mathcal{S}\times\mathcal{S}}\left(\frac{1}{|s^1|}{\sum_{\substack{i\\s^1_i\equiv s^2_i}}1}\right) &&& \text{Seq. identity}
\end{align}
where $s_i$ is the nucleotide at position $i$ in sequence $s$. Intuitively, this measure captures the diversity of sequences generated by a given method. Next, the Boltzmann frequency is defined, for a structure $\Struct$ and a sequence $s$ as:
\begin{align}
  e^{\frac{-\ES(s,\Struct)}{RT}}/\mathcal{Z}^s  &&&\text{Frequency}
\end{align}
where $\mathcal{Z}^s$ is the partition function of sequence $s$. This measure tells us how dominant is a structure $\Struct$ in the Boltzmann ensemble of structure over a sequence $s$. A high value implies a stable structure. We compute this frequency with \rnafold from the \texttt{Vienna Package 2.0} \citep{Hofacker:1994}.
}

@@

The thermodynamical stability of the target structure on the designed sequence is another important property when estimating the performance of RNA design algorithms. We estimate the quality of our solutions in Figure~\ref{fig:freq}. First, we observe a slow decline of the structure stability (i.e. the frequency) when the target structure increases in size. Yet, for an average \GCContent, the frequency stays over $10\%$ even at size of $100$ nucleotides. Next, we note that for the most difficult target structures (i.e. the longer ones or those with high percentages of unpaired nucleotides in loops) the \GCContent have a limited (almost null) influence on the stability of the target structure on the designed sequence. By contrast, this less true for easiest and small structures with only few free nucleotides in internal loops.
}

>>>>

Figure~\ref{fig:time} presents the average times spent running \ourprog+\RNAinverse to generate one sequence
with the required \GCContent. As expected, the time grows linearly
in function of the length of the structures for \ourprog.  
%A highly intriguing feature is the substantial decrease of the time 
%required for running \RNAinverse when $15\%$ of nucleotides are unpaired in the target structure.
%However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).

@@

%As discussed in Sec.~\ref{sec:implementation}, we noticed a highly decreased computational time needed to generate the sequences with $15\%$ free 
%nucleotides in the loops. 
We noticed a highly decreased structural sensitivity for the sequences with $15\%$ free nucleotides in the loops. 
%A highly intriguing feature is the substantial decrease of the time 
%required for running \RNAinverse when $15\%$ of nucleotides are unpaired in the target structure.
However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).

@@

The sequence identity is defined over a set $\mathcal{S}$ of aligned sequences (in our case, all sequences have the same length and can be trivially aligned) as :
\begin{align}
  \sum_{s^1,s^2 \in\mathcal{S}\times\mathcal{S}}\left(\frac{1}{|s^1|}{\sum_{\substack{i\\s^1_i\equiv s^2_i}}1}\right) &&& \text{Seq. identity}
\end{align}
where $s_i$ is the nucleotide at position $i$ in sequence $s$. Intuitively, this measure captures the diversity of sequences generated by a given method. Next, the Boltzmann frequency is defined, for a structure $\Struct$ and a sequence $s$ as:
\begin{align}
  e^{\frac{-\ES(s,\Struct)}{RT}}/\mathcal{Z}^s  &&&\text{Frequency}
\end{align}
where $\mathcal{Z}^s$ is the partition function of sequence $s$. This measure tells us how dominant is a structure $\Struct$ in the Boltzmann ensemble of structures over a sequence $s$. A high value implies a stable structure. We compute this frequency with \rnafold from the \texttt{Vienna Package 2.0} \citep{Hofacker:1994}.
}

@@

The thermodynamical stability of the target structure on the designed sequence is another important property when estimating the performance of RNA design algorithms. We estimate the quality of our solutions in Figure~\ref{fig:freq}. First, we observe a slow decline of the structure stability (i.e. the frequency) when the target structure increases in size. Yet, for an average \GCContent, the frequency stays over $10\%$ even at size of $100$ nucleotides. Next, we note that for the most difficult target structures (i.e. the longer ones or those with high percentages of unpaired nucleotides in loops) the \GCContent have a limited (almost null) influence on the stability of the target structure on the designed sequence. By contrast, this is less true for easiest and small structures with only few free nucleotides in internal loops.
}
